---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sdmtools

<!-- badges: start -->
[![R-CMD-check](https://github.com/idem-lab/sdmtools/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/idem-lab/sdmtools/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/idem-lab/sdmtools/branch/main/graph/badge.svg)](https://app.codecov.io/gh/idem-lab/sdmtools?branch=main)
![GitHub License](https://img.shields.io/github/license/idem-lab/sdmtools)
[![Lifecycle:](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
![GitHub commit activity](https://img.shields.io/github/commit-activity/w/idem-lab/sdmtools)
![GitHub Downloads (all assets, all releases)](https://img.shields.io/github/downloads/idem-lab/sdmtools/total)
![GitHub last commit](https://img.shields.io/github/last-commit/idem-lab/sdmtools)
![GitHub R package version](https://img.shields.io/github/r-package/v/idem-lab/sdmtools)
![GitHub commits since latest release](https://img.shields.io/github/commits-since/idem-lab/sdmtools/latest)
<!-- badges: end -->

A set of helper functions to facilitate species distribution modelling.

## Installation

You can install  `sdmtools` with:

``` r
install.packages(
"sdmtools",
repos = "https://idem-lab.r-universe.dev"
)
```

## Data
```{r}
library(sdmtools)
```


`raster_to_terra` — an annotated equivalence table of functions from the `raster` and `terra`. First 5 lines:

```{r, echo=FALSE}
knitr::kable(
  raster_to_terra[1:5,],
  format = "html",
  col.names = c("raster", "terra", "comment for terra use"),
  column.spec = c("c", "c", ">{\\raggedright}p{8cm}")
)

```


 `global_regions` — a tibble showing the WHO region,  UN region, and continent for for 249 countries and country-like things. First 5 lines:
 
```{r, echo=FALSE}
knitr::kable(
  global_regions[1:5,],
  format = "html",
  #col.names = c("raster", "terra", "comment for terra use"),
  column.spec = c("c", "c", ">{\\raggedright}p{8cm}")
)

```


### Data-generating functions

The package `terra` is fiddly about storing its `spat...` objects in packages, so we chose to generate example spatial data on-demand using functions, rather than storing it.

`example_raster` — an example `spatRaster`.

```{r}
library(terra)
r <- example_raster()
r
plot(r)
```

`example_vector` — an example `spatVector`.

```{r}
library(terra)
v <- example_vector()
v
plot(v)
```

`make_africa_mask` — makes a mask layer of Africa based on shapefiles from `malariaAtlas::getShp`.
Can produce either a `SpatRaster` or `SpatVector`.


```{r}
library(terra)
africa_mask <- make_africa_mask(type = "vector")
plot(africa_mask)
```

## Function examples

`rastpointplot` — simple utility to plot a raster with points over it.

```{r}
rastpointplot(r,v)
```

`source_R` — source all R files in a target directory
```r
source_R("/Users/frankenstein/project/R") # do not run
```


`import_rasts` — import all rasters from a directory into a single object
```r
rasters <- import_rasts("/data/grids/covariates") # do not run
```


`split_rast` — split a raster.
```{r}
r <- example_raster()

s <- split_rast(r, grain = 2)

s
```
```{r}
ps <- lapply(
  s,
  FUN = extend,
  y = r
) |>
  rast()

c(
  r,
  ps
) |>
  plot()

```


### Functions for a species distribution modelling workflow

We have some covariate layers: `cov1` and `cov2`
```{r}
library(terra)

cov1 <- example_raster(
  seed = -44,
  layername = "cov1"
)
cov2 <- example_raster(
  seed = 15.3,
  layername = "cov2"
)
```

`standardise_rast` — standardise a `spatRaster` by transforming it to have a mean of zero and standard deviation of one
```{r}
cov1_st <- standardise_rast(cov1)
cov2_st <- standardise_rast(cov2)

covs <- c(cov1_st, cov2_st)

covs

plot(covs)
```


We have some presences and absences
```{r}
presences <- example_vector(seed = 68) %>%
  as.data.frame(geom = "xy")
absences <- example_vector(seed = 9.6) %>%
  as.data.frame(geom = "xy")

presences
```


`extract_covariates` — extract covariate values from `spatRaster` or `raster` layers for a given set of points

Pass in either `presences` and `absences` as a `data.frame` or `tibble` of with , or `presences_and_absences` as a single data frame points with a presence or ID column(s)
```{r}
sdm_data <- extract_covariates(
  covariates = covs,
  presences = presences,
  absences = absences
)
```

We can then make a spatial prediction of our model using `predict_sdm` and write and read it out in a single step with `writereadrast`, and write it to a temporary file with `temptif`:
```{r}
# first we make a simple model, using data from above
m <- glm(presence ~ cov1 + cov2, data = sdm_data)

prediction_rast <- predict_sdm(m, covs) |>
  writereadrast(filename = temptif())

plot(prediction_rast)
```



