---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sdmtools

<!-- badges: start -->
[![Licence](https://img.shields.io/github/license/Ileriayo/markdown-badges?style=for-the-badge)](./LICENSE)
<!-- badges: end -->

`sdmtools` is a set of helper functions to facilitate species distribution modelling.

## Installation

You can install the development version of sdmtools from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("idem-lab/sdmtools")
```

## Data
```{r}
library(sdmtools)
```


`raster_to_terra` an annotated equivalence table of functions from the `raster` and `terra` packages incorporating section **XXVIII New method names** from `terra`'s [package notes](https://cran.r-project.org/web/packages/terra/terra.pdf) 
```{r, echo=FALSE}
knitr::kable(
  raster_to_terra,
  format = "html",
  col.names = c("raster", "terra", "comment for terra use"),
  column.spec = c("c", "c", ">{\\raggedright}p{8cm}")
)

```


### Data-generating functions

`example_raster` an example `spatRaster`
```{r}
library(terra)
r <- example_raster()
plot(r)
```

`example_vector` an example `spatVector`
```{r}
library(terra)
v <- example_vector()
plot(v)
```

## Function examples

`source_R` — source all R files in a target directory
```r
source_R("/Users/frankenstein/project/R")
```




`import_rasts` — Import all rasters from a directory into a single object
```r
rasters <- import_rasts("/data/grids/covariates")
```

*A species distribution modelling workflow:*

We have some covariate layers: `cov1` and `cov2`
```{r}
library(terra)

cov1 <- example_raster(seed = -44)
cov2 <- example_raster(seed = 15.3)

names(cov1) <- "cov1"
names(cov2) <- "cov2"

covs <- c(cov1, cov2)

plot(covs)
```

We have some presences and absences
```{r}
presences <- example_vector(seed = 68) %>%
  as.data.frame(geom = "xy")
absences <- example_vector(seed = 9.6) %>%
  as.data.frame(geom = "xy")
```


`extract_covariates` — extract covariate values from `spatRaster` or `raster` layers for a given set of points
```{r}

extract_covariates(
  covariates = covs,
  presences = presences,
  absences = absences
)
```

`predict_sdm` — made a spatial prediction from a species distribution model and covariate layers
```{r}



m <- glm(z ~ cov1 + cov2, data = sdm_data)

prediction <- predict_sdm(m, covs)
```


`rastpointplot` — simple utility to plot a raster with points over it
```{r}
r <- example_raster()
v <- example_vector()
rastpointplot(r,v)
```
