---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sdmtools

<!-- badges: start -->
[![Licence](https://img.shields.io/github/license/Ileriayo/markdown-badges?style=for-the-badge)](./LICENSE)
<!-- badges: end -->

`sdmtools` is a set of helper functions to facilitate species distribution modelling.

## Installation

You can install the development version of sdmtools from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("geryan/sdmtools")
```

## Data
```{r}
library(sdmtools)
```


`raster_to_terra` an annotated equivalence table of functions from the `raster` and `terra` packages
```{r}
raster_to_terra
```


`r` an example `spatRaster`
```r
r
plot(r)
```

`v` an example `spatVector`
```r
v
plot(v)
```

## Function examples

`source_R` — source all R files in a target directory
```r
source_R("/Users/frankenstein/project/R")
```



`predict_sdm` — made a spatial prediction from a species distribution model and covariate layers
```r
m <- glm(z ~ cov1, cov2, data = sdm_data)

prediction <- predict_sdm(m, covs)
```

`import_rasts` — Import all rasters from a directory into a single object
```r
rasters <- import_rasts("/data/grids/covariates")
```

`rastpointplot` — plot a raster with points over it
```r
rastpointplot(r,v)
```

`extract_covariates` — extract covariate values from `spatRaster` or `raster` layers for a given set of points
```r
# example
```


